---
title: "FARS"
author: "Nivea Bekal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FARS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The `FARS` package contains functions provided for the students. These functions were written to analyze data from the [US National Highway Traffic Safety Administration's Fatality Analysis Reporting System](https://www.nhtsa.gov) (FARS). FARS is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

## Dependencies

You need to load the following packages before starting:

```{r, echo=TRUE, include=TRUE}
install.packages("devtools")
library(devtools)
library(dplyr)
library(maps)
library(readr)
library(tidyr)
```


## Functions

```{r}

devtools::install_github("https://github.com/niveabekal04/farsReport")
library(farsReport)
farsReport::fars_summarize_years(years = c(2013, 2014))
```

### FARS mapplot per state
The `fars_map_state(state.num, year)` function will plot FARS accidents for specififed `year` on border map of the state (specified by `state.num` parameter). This function does not return any data, the side effect of the function is the plot of FARS accident on the border map of the state.

```{r, fig.show='hold'}
library(farsReport)
farsReport::fars_map_state(state.num = 13, year = 2014)
farsReport::fars_map_state(state.num = 34, year = 2014)
```

## FARS data structure
The function responsible for reading the data is `fars_read(filename)`, whereas filename argument is generated via the `fars_make_filename(year)` function, which returns path to a `year` dataset from saved range of years 2013-2015 (with default `year` being 2013).

The structure of an example dataset (for year 2013) is written below:

```{r, echo=FALSE}
library(farsReport)
knitr::kable(str(farsReport::fars_read(farsReport::fars_make_filename())))
```

